An Efficient Feedback Calibration Algorithm for Direct Imaging Radio Telescopes

### 2016-06-08 cal_demo/ms.tex
more edits... updated sim images figure, changed model for data

Images created at the beginning of calibration and at the end are shown in 
figure~\ref{fig:sim_images}. Each image is obtained over 10~ms integration, corresponding to 
all snapshot images created with a given set of gain estimates. The top panel shows the image 
produced with our initialized unity gains. Because the phases are completely random, the 
image is essentially noise with the primary beam evident. After 20 iterations, the image is far 
more clear, shown in the bottom panel. Each of the ten simulated sources is clearly visible, 
indicated with red circles.

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{fig5a.pdf}
\includegraphics[width=\columnwidth]{fig5b.pdf}
\caption{Images formed during simulated calibration. \emph{Top:} An image generated from a 
single frequency channel and 10 ms integration after the gain estimates are randomized. As 
expected with random phases, the image is completely noisy, with the shape of the primary 
beam evident. \emph{Bottom:} An image formed after calibration, again with a single frequency 
channel and 10 ms integration. The ten simulated (and modeled) point sources are easily 
visible, and highlighted with red circles.
}
\label{fig:sim_images}
\end{center}
\end{figure}

@@

We proceed by forming model visibilities. We model only two bright objects as point sources: 
Cyg A with flux 16611.68 Jy \citep{coh07}; and Cas A with flux 17693.9 Jy \citep{kas07}. 
Because the raw data is attenuated by the primary beam of the instrument, we also account for 
this in our model using beam values consistent with \cite{hic12}.

@@

Figure~\ref{fig:data_images} shows the improvement in the images due to our calibration. The 
left panel shows the uncalibrated image integrated over 51.2 ms, 29 kHz. Cyg~A is prominent 
near the center of the image, and Cas~A is also clearly visible in the upper right. The middle 
panel shows the image produced after calibration with identical integration time and bandwidth. 
The sidelobes throughout the image are significantly suppressed, and the galactic plane is 
much more evident, despite only modeling Cyg~A and Cas~A. We also note that the feature 
just to the right of Cyg~A is dimmer in the calibrated image, better matching the expected flux 
from the global sky model (GSM, \citealt{deo08}). We show the GSM \citep{pri16} modulated 
by two factors of the LWA power pattern in the right panel of figure~\ref{fig:data_images} for 
reference.

\begin{figure*}
\begin{center}
\includegraphics[width=0.3\linewidth]{fig8a.pdf}
\includegraphics[width=0.3\linewidth]{fig8b.pdf}
\includegraphics[width=0.3\linewidth]{fig8c.pdf}
\caption{Images produced before (left) and after (middle) calibrating LWA data. These images 
were produced with 29 kHz bandwidth and 51.2 ms integration. The calibrated image shows 
significant reduction in sidelobe rumble throughout, while retaining the prominent Cyg~A and 
Cas~A sources. The galactic plane is also substantially more evident after calibration. For 
reference, the GSM is shown in the same coordinates and weighted by two factors of the LWA 
primary beam in the right panel. While much of the GSM is visible in the calibrated image, we 
note that the model used to calibrate was only the two bright sources, Cyg~A and Cas~A.
}
\label{fig:data_images}
\end{center}
\end{figure*}

>>>>

Images created at the beginning of calibration and at the end are shown in 
figure~\ref{fig:sim_images}. Each image is obtained over 10~ms integration, corresponding to 
all snapshot images created with a given set of gain estimates. The left panel shows the image 
produced with our initialized unity gains. Because the phases are completely random, the 
image is essentially noise with the primary beam evident. After 20 iterations, the image is far 
more clear, shown in the middle panel. 
The ten sources are visible, along with rumble throughout the image. By comparing with
a perfectly calibrated image (right panel), we see that the rumble and the shapes of the
point sources are matched between the two images. We therefore conclude that the 
artifacts visible are dominated by the point spread function, rather than calibration errors.

\begin{figure*}
\begin{center}
\includegraphics[width=.3\linewidth]{fig5a.pdf}
\includegraphics[width=.3\linewidth]{fig5b.pdf}
\includegraphics[width=.3\linewidth]{fig5c.pdf}
\caption{Images formed during simulated calibration. \emph{Left:} An image generated from a 
single frequency channel and 10 ms integration after the gain estimates are randomized. As 
expected with random phases, the image is completely noisy, with the shape of the primary 
beam evident. \emph{Middle:} An image formed after calibration, again with a single frequency 
channel and 10 ms integration. Now the simulated point sources are visible in the image. 
\emph{Right:} An image generated from the same simulated data as the middle image, but 
with perfect calibration correction. Many of the features seen in the middle image are also
seen here, confirming that they are point spread function artifacts. The ten simulated (and modeled) point sources are highlighted with red circles.
\label{fig:sim_images}
}
\end{center}
\end{figure*}

@@

We proceed by forming model visibilities. We model only two bright objects as point sources: 
Cyg A with flux 20,539 Jy, and Cas A with flux 19,328 Jy \citep{lan12}. 
Because the raw data is attenuated by the primary beam of the instrument, we also account for 
this in our model using beam values consistent with \cite{hic12}.

@@

Figure~\ref{fig:data_images} shows the improvement in the images due to our calibration. The 
left panel shows the uncalibrated image integrated over 51.2 ms, 29 kHz. Cyg~A is prominent 
near the center of the image, and Cas~A is also clearly visible in the upper right. The middle 
panel shows the image produced after calibration with identical integration time and bandwidth. 
The sidelobes throughout the image are significantly suppressed, and the galactic plane is 
much more evident, despite only modeling Cyg~A and Cas~A. We also note that the feature 
just to the right of Cyg~A is dimmer in the calibrated image, better matching the expected flux 
from the global sky model (GSM, \citealt{deo08}). 
For reference we should the GSM \citep{pri16} in the right panel of 
Figure~\ref{fig:data_images}, convolved by the LWA point spread function and weighted
by  two factors of the primary beam to match the data images.

\begin{figure*}
\begin{center}
\includegraphics[width=0.3\linewidth]{fig8a.pdf}
\includegraphics[width=0.3\linewidth]{fig8b.pdf}
\includegraphics[width=0.3\linewidth]{fig8c.pdf}
\caption{Images produced before (left) and after (middle) calibrating LWA data. These images 
were produced with 29 kHz bandwidth and 51.2 ms integration. The calibrated image shows 
significant reduction in rumble throughout, while retaining the prominent Cyg~A and 
Cas~A sources. The galactic plane is also substantially more evident after calibration. For 
reference, the GSM is shown in the 
right panel convolved with the LWA point spread function and weighted by two factors
of the primary beam to place it in the same holographic frame as the data images. While much of the GSM is visible in the calibrated image, we 
note that the model used to calibrate was only the two bright sources, Cyg~A and Cas~A.
}
\label{fig:data_images}
\end{center}
\end{figure*}
